### 2023-02-23 manuscript.Rmd
Add rating cosine results

# Specify directory paths
data_dir <- here("data")
files_dir <- here("manuscript_files")
output_dir <- here("output")
report_dir <- here("output/qc_reports")

@@

# Re-run steps that take a long time to compute?
run <- list(
  eeg_processing = FALSE,
  mixed_models = FALSE,
  figures = FALSE
)

@@

For each unfamiliar object, a pair of German keywords (a noun and a verb) was selected, describing the typical function or use of the object in a way that could be related to its visual features and their configuration (e.g., *Stromstärke, messen* [electric current, measuring]; *Tonpott, trommeln* [clay pot, drumming]; [see Supplementary Table 1]{color="blue"}).
As our central experimental manipulation, half of the objects were presented together with keywords that correctly matched their respective function, whereas the other half of the objects were presented together with non-matching keywords belonging to one of the other objects.
The matching keywords were expected to induce semantically informed perception, that is, participants suddenly understanding what kind of object they were viewing.
The non-matching keywords were expected to prevent such an understanding and keep the perception of the object semantically [uninformed]{color="blue"}.
All participants saw each unfamiliar object with only one type of keywords (matching or non-matching).
This assignment of keywords to objects was counterbalanced across participants so that each object was presented with matching keywords (leading to semantically informed perception) and non-matching keywords (leading to [uninformed]{color="blue"} perception) to an equal number of participants.
The experiment was programmed and displayed using Presentation® software (Neurobehavioral Systems, Inc., Berkeley, CA, www.neurobs.com).

@@

The experiment consisted of three phases (see Figure 1A).
In the *pre-insight* phase, after written informed consent had been obtained and the EEG had been prepared, all 240 familiar and unfamiliar objects were presented once in random order and without any keywords.
Each trial consisted of a fixation cross presented in the middle of the screen for 0.5 s, followed by the presentation of the object until participants made a response or until a time out after 3 s.
The inter-trial interval was 0.5 s and participants took a self-timed break after each block of 60 objects.
The task, which was kept the same across all three phases of the experiment, was to classify each object using one of four response alternatives:
(a) *Ich weiß, was das ist, oder habe eine starke Vermutung* [I know what this is or have a strong assumption],
(b) *Ich habe eher eine Vermutung, was das ist* [I rather have an assumption what this is],
(c) *Ich habe eher keine Vermutung, was das ist* [I rather have no assumption what this is], or
(d) *Ich weiß nicht, was das ist, und habe auch keine Vermutung* [I don’t know what this is and have no assumption].
Participants were asked to respond as quickly and as accurately as possible by pressing one out of four buttons with the index or middle finger of their left or right hand, respectively.
The mapping of the rating scale to the four buttons (left to right or right to left) was counterbalanced across participants.

@@

>>>>

# Specify directory paths
data_dir <- here("data")
rating_dir <- here(data_dir, "ratings")
files_dir <- here("manuscript_files")
output_dir <- here("output")
report_dir <- here(output_dir, "qc_reports")

@@

# Re-run steps that take a long time to compute?
run <- list(
  rating_semantic_analysis = FALSE,
  eeg_processing = FALSE,
  mixed_models = FALSE,
  figures = FALSE
)

@@

For each unfamiliar object, we created a pair of German keywords (a noun and a verb), describing the typical function or use of the object in a way that could be related to its visual features and their configuration (e.g., *Stromstärke, messen* [electric current, measuring]; *Tonpott, trommeln* [clay pot, drumming]; [see Supplementary Table 1]{color="blue"}).
As our central experimental manipulation, half of the objects were presented together with keywords that correctly matched their respective function, whereas the other half of the objects were presented together with non-matching keywords belonging to one of the other objects.
The matching keywords were expected to induce semantically informed perception, that is, participants suddenly understanding what kind of object they were viewing.
The non-matching keywords were expected to prevent such an understanding and keep the perception of the object semantically [uninformed]{color="blue"}.
All participants saw each unfamiliar object with only one type of keywords (matching or non-matching).
This assignment of keywords to objects was counterbalanced across participants so that each object was presented with matching keywords (leading to semantically informed perception) and non-matching keywords (leading to [uninformed]{color="blue"} perception) to an equal number of participants.
The experiment was programmed and displayed using Presentation® software (Neurobehavioral Systems, Inc., Berkeley, CA, www.neurobs.com).

```{r, ratings}
# Read table of object stimuli + keywords for the main experiment
keywords <- read_csv(
  here("materials", "keywords.csv"),
  col_types = cols(
    familiarity = col_factor(levels = c("familiar", "unfamiliar")),
    item_id = col_factor()
  )
) %>%
  mutate(
    match = str_c(match_noun, ", ", match_verb),
    match_en = str_c("{[", match_noun_en, ", ", match_verb_en, "]}"),
    nonmatch = str_c(nonmatch_noun, ", ", nonmatch_verb),
    nonmatch_en = str_c("{[", nonmatch_noun_en, ", ", nonmatch_verb_en, "]}"),
  )

# (Re-)do latent semantic analysis of online rating study data
rating_file <- here(output_dir, "rating.csv")
if (run$rating_semantic_analysis) {
  # Load pre-trained German word2vec embedding
  word2vec <- load_word2vec(rating_dir)

  # Read online rating study data
  rating_raw <- read_csv(
    here(data_dir, "ratings", "rating_data.csv"),
    col_types = cols(), locale = locale(encoding = "latin1")
  )
  rating_substantives <- select(rating_raw, S001_01:S242_01)
  rating_verbs <- select(rating_raw, V001_01:V242_01)

  # Prepare a copy of the keywords to add to the rating study data
  rating_n <- nrow(rating_raw)
  keywords_for_rating <- keywords %>%
    arrange(familiarity, filename) %>%
    select(item_id, familiarity, filename, match, nonmatch) %>%
    replicate(rating_n, ., simplify = FALSE) %>%
    bind_rows()

  # Compute cosine similarities between rating study responses and keywords
  rating <-
    map2_dfc(rating_substantives, rating_verbs, str_c, sep = ", ") %>%
    mutate(participant_id = rating_raw$CASE) %>%
    pivot_longer(
      -participant_id,
      names_to = "question_id", values_to = "response"
    ) %>%
    bind_cols(keywords_for_rating, .) %>%
    pivot_longer(
      c(match, nonmatch),
      names_to = "condition", values_to = "keywords"
    ) %>%
    mutate(across(c(keywords, response), str_to_lower)) %>%
    mutate(cosine = compute_cosine(keywords, response, word2vec))

  # Save to file
  write_csv(rating, rating_file)
}

# Read the rating study results incl. cosine similarities
rating <- read_csv(
  rating_file,
  col_types = cols(item_id = col_factor(), participant_id = col_factor())
) %>%
  filter(!is.na(cosine)) %>%
  mutate(
    fam_condition = paste(familiarity, condition) %>%
      factor(
        levels = c("familiar match", "unfamiliar match", "unfamiliar nonmatch"),
        labels = c("familiar", "matching", "nonmatching")
      )
  )

# Compare cosine similarities between object conditions
contrasts(rating$fam_condition) <- contr.sdif(3)
model_rating <- lmer(
  cosine ~ fam_condition +
    (fam_condition | participant_id) +
    (1 | item_id),
  data = rating,
  control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e6))
)
(summary_rating <- summary(model_rating))

# Compute condition means and pairwise contrasts
emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
emm_rating <- emmeans(
  model_rating,
  specs = consec ~ fam_condition,
  rev = TRUE,
  adjust = "bonferroni"
)
rating_means <- as.data.frame(emm_rating$emmeans)
rating_contrasts <- as.data.frame(emm_rating$contrasts) %>%
  rename(t_value = `t.ratio`, p_value = `p.value`)

# rating_pairwise <- rating$participant_id %>%
#   unique() %>%
#   combn(m = 2) %>%
#   data.frame() %>%
#   map_dfr(function(pair) {
#     id_1 <- pair[1]
#     id_2 <- pair[2]
#     response_1 <- filter(rating, participant_id == id_1)$response
#     response_2 <- filter(rating, participant_id == id_2)$response
#     tibble(
#       item_id = keywords$item_id,
#       familiarity = keywords$familiarity,
#       filename = keywords$filename,
#       pair_id = str_c(id_1, id_2, sep = "_"),
#       participant_id_1 = id_1,
#       participant_id_2 = id_2,
#       response_1,
#       response_2,
#       cosine = compute_cosine(response_1, response_2, word2vec)
#     )
#   })

# write_csv(rating_pairwise, here(output_dir, "rating_pairwise.csv"))

# mod_pairwise <- lmer(
#   cosine ~ familiarity + (familiarity | pair_id) + (1 | filename),
#   data = rating_pairwise
# )
# (summary_pairwise <- summary(mod_pairwise))

# Rmisc::summarySEwithin(
#   data = rating_pairwise, measurevar = "cosine", withinvars = "familiarity",
#   idvar = "pair_id", na.rm = TRUE
# ) %>%
#   ggplot(aes(x = familiarity, y = cosine, color = familiarity)) +
#   geom_violin(data = rating_pairwise) +
#   coord_cartesian(ylim = c(0.0, 1.0)) +
#   geom_pointrange(aes(ymin = cosine - se, ymax = cosine + se)) +
#   theme_classic()
```

As a manipulation check, we ran an online rating study where we presented 10 German speakers (3 female, 7 male, mean age = 25.3 years, range 20--35 years) with all 240 visual objects in random order and asked them to generate their own keywords that would describe the presumed function of the object.
We used latent semantic analysis with a word2vec embedding pre-trained on the German Wikipedia (deepset GmbH, Berlin, Germany, www.deepset.ai/german-word-embeddings) to estimate semantic distances between these participant-generated keywords and the keywords used in our experiment (Supplementary Methods; Supplementary Figures 1 and 2).
In brief, it was substantially easier for participants to come with correct descriptions of familiar objects (mean cosine distance ± SE = `r filter(rating_means, fam_condition == "familiar")$emmean` ± `r filter(rating_means, fam_condition == "familiar")$SE`) than for unfamiliar objects (`r filter(rating_means, fam_condition == "matching")$emmean` ± `r filter(rating_means, fam_condition == "matching")$SE`), $t$(`r print_dof(filter(rating_contrasts, contrast == "familiar - matching")$df)`) = `r filter(rating_contrasts, contrast == "familiar - matching")$t_value`, $p_\text{corr}$ `r print_p(filter(rating_contrasts, contrast == "familiar - matching")$p_value)` (linear mixed-effects model).
In fact, this similarity between participant-generated keywords for the unfamiliar objects and the object-matching keywords presented in the main experiment was only slightly higher than the similarity between participant-generated keywords and the object-non-matching keywords presented in the main experiment (`r filter(rating_means, fam_condition == "nonmatching")$emmean` ± `r filter(rating_means, fam_condition == "nonmatching")$SE`), $t$(`r print_dof(filter(rating_contrasts, contrast == "matching - nonmatching")$df)`) = `r filter(rating_contrasts, contrast == "matching - nonmatching")$t_value`, $p_\text{corr}$ `r filter(rating_contrasts, contrast == "matching - nonmatching")$p_value` (linear mixed-effects model).

@@

The main EEG experiment consisted of three phases (see Figure 1A).
In the *pre-insight* phase, after written informed consent had been obtained and the EEG had been prepared, all 240 familiar and unfamiliar objects were presented once in random order and without any keywords.
Each trial consisted of a fixation cross presented in the middle of the screen for 0.5 s, followed by the presentation of the object until participants made a response or until a time out after 3 s.
The inter-trial interval was 0.5 s and participants took a self-timed break after each block of 60 objects.
The task, which was kept the same across all three phases of the experiment, was to classify each object using one of four response alternatives:
(a) *Ich weiß, was das ist, oder habe eine starke Vermutung* [I know what this is or have a strong assumption],
(b) *Ich habe eher eine Vermutung, was das ist* [I rather have an assumption what this is],
(c) *Ich habe eher keine Vermutung, was das ist* [I rather have no assumption what this is], or
(d) *Ich weiß nicht, was das ist, und habe auch keine Vermutung* [I don’t know what this is and have no assumption].
Participants were asked to respond as quickly and as accurately as possible by pressing one out of four buttons with the index or middle finger of their left or right hand, respectively.
The mapping of the rating scale to the four buttons (left to right or right to left) was counterbalanced across participants.

@@

```{r, ratingfigs}
rating %>%
  filter(!is.na(cosine)) %>%
  mutate(condition_comb = paste(familiarity, condition)) %>%
  ggplot(aes(x = condition_comb, y = cosine, color = condition_comb)) +
  geom_violin() +
  geom_boxplot(width = 0.2) +
  scale_color_manual(values = c("#984ea3", "#e41a1c", "#377eb8")) +
  labs(x = NULL, y = "Cosine distance") +
  coord_cartesian(ylim = c(0.0, 1.0)) +
  theme_classic() +
  theme(legend.position = "none")

rating %>%
  filter(
    !is.na(cosine) & familiarity == "unfamiliar" & condition == "match"
  ) %>%
  select(item_id, condition, cosine) %>%
  group_by(item_id, condition) %>%
  summarize(
    mean = mean(cosine),
    sd = sd(cosine),
    n = n(),
    se = sd / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(item_id = fct_reorder(item_id, desc(mean))) %>%
  # filter(mean > 0.7) %>%
  ggplot(
    aes(
      x = item_id, y = mean, color = condition,
      ymin = mean - se, ymax = mean + se
    )
  ) +
  geom_pointrange() +
  labs(x = "Stimulus ID", y = "Cosine distance to keywords") +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 90),
    legend.position = "none"
  )
```
